A new machine-learning-based prediction of survival in patients with end-stage liver disease

### 2022-12-19 analysis/article.Rmd
feat: add some paragraphs for variable importance

@@

@@

```{r selvar, echo = FALSE}
bsrank <- as.vector(rank(-ameld:::.selvars(bootrcv$models, bootrcv$s), ties = "first")["NA_S"])

bsranktxt <-
    if (!is.na(bsrank)) {
        paste0("in bootstrap selection rank ", bsrank, " of ", ncol(amelddata$x))
    } else {
        paste0("it was never selected in any boostrap sample")
    }

rfrank <- as.vector(rank(-rngr$variable.importance, ties = "first")["NA_S"])
rfranktxt <-
    paste0("in a random forest variable importance rank ", rfrank, " of ", ncol(amelddata$x))
```

Although the Organ Procurement and Transplantation Network recommends the use
of MELD-Na over MELD, the sodium does not add any information to the score.
Neither in the bootstrap selection nor in the variable importance of a
random forest model does the sodium gain any important weight
(`r bsranktxt` and `r rfranktxt`, see \@ref(fig:vimpbs), \@ref(fig:vimprf)).
In addition, sodium levels can be affected by diuretic therapy, and
the discrimination performance of MELD-Na has already been shown by
@kartoun2017 to be insignificantly different from the classical MELD.
Consequently it is not a component of our model.

>>>>

Further we determined the variable importance by counting how often the
variable was chosen during the feature selection in the bootstrap process
described above and by the logrank value returned by one random forest
implementation [@R-ranger].

@@

## Variable importance

```{r selvar, echo = FALSE}
bsrank <- rank(-ameld:::.selvars(bootrcv$models, bootrcv$s), ties = "first")

bsrankSodium <- as.vector(bsrank["NA_S"])

bsranktxtSodium <-
    if (!is.na(bsrankSodium)) {
        paste0(
            "in bootstrap selection rank ", bsrankSodium,
            " of ", ncol(amelddata$x)
        )
    } else {
        paste0("it was never selected in any boostrap sample")
    }

rfrank <- rank(-rngr$variable.importance, ties = "first")
rfrankSodium <- as.vector(rfrank["NA_S"])
rfranktxtSodium <- paste0(
    "in a random forest variable importance rank ", rfrankSodium, " of ",
    ncol(amelddata$x)
)
```

The AMELD selected features
`r enum(gsub("^([A-Z])([a-z])", "\\L\\1\\2", cfnms, perl = TRUE)) `
are the `r max(bsrank[names(cfnms)])` highest ranked features in the bootstrap
process. In the random forest logrank they are amoung the
`r max(rfrank[names(cfnms)])` highest ranked features out of `r length(rfrank)`
features.
Sodium and albumin introduced by MELD-Na and MELD 3.0 are on position
`r paste(bsrank[c("NA_S", "ALB_S")], collapse = " and ")` for the bootstrap and
`r paste(rfrank[c("NA_S", "ALB_S")], collapse = " and ")` for the random forest
variable importance ranking, respectively
(see \@ref(fig:vimpbs), \@ref(fig:vimprf) for the complete rankings).

@@

Although the Organ Procurement and Transplantation Network recommends the use
of MELD-Na over MELD, the sodium does not add any information to the score.
Neither in the bootstrap selection nor in the variable importance of a
random forest model does the sodium gain any important weight
(`r bsranktxtSodium` and `r rfranktxtSodium`, see \@ref(fig:vimpbs), \@ref(fig:vimprf)).
In addition, sodium levels can be affected by diuretic therapy, and
the discrimination performance of MELD-Na has already been shown by
@kartoun2017 to be insignificantly different from the classical MELD.
Consequently it is not a component of our model.

The albumin included in MELD 3.0 is part of the total protein that has
throughout higher rankings in both variable importance measurements
(`r bsrank["ALB_S"]` vs. `r bsrank["PROT_S"]` in the bootstrap selection
and `r rfrank["ALB_S"]` vs. `r rfrank["PROT_S"]` in the random forest variable
importance, see \@ref(fig:vimpbs), \@ref(fig:vimprf)).
Using total protein instead of albumin as synthesis marker
offers more information about additional globulins and especially
coagulation factors that are incompletely measured by INR.
